---
title: "Lab10_inclass"
author: "Qianqian Tao"
format: pdf
---

ADK (Adenelate Kinase) important drug target, know its molecular mechanism

We will begin with getting an example ADK sequence from the database.
We will then use this to find all 
```{r}
library(bio3d)

aa <- get.seq("1ake_A")
aa
```
We can now run BLAST with this sequence
```{r}
#b <- blast.pdb(aa)
```

```{r}
#same as plot.blast(b)
#hits <- plot(b)
```
Let's see what is in our `hits` object

```{r}
#hits$pdb.id
```

Now we can download all these top hits PDB structure files:
```{r}
# Download releated PDB files
#?get.pdb
```
```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

Now I want to align and supperpose these structures which are all over the place

```{r}
# Align releated PDBs; fit on top of each other, exefile run alignment on the same package 
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

Let's have a look at our `pdbs` object
```{r}
pdbs
```

We have our aligned and superposed structures we can perform all sorts of analysis on them. Let's do PCA...
```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```

Results of PCA on Adenylate kinase X-ray structures. Each dot represents one PDB structure.

We can cluster the structures by RMSD (or any other method). 
```{r}
rd <- rmsd(pdbs)
# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col=grps)
#plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```
We can make a wee movie - a trajectory of the major differences of 
```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```

