---
title: "Lab11_inclass"
author: "Qianqian Tao"
date: "2023-05-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 1. Proportion og G/G in a population

Downloaded a CSV file from Ensemble <
https://useast.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39815101-39975102;v=rs8067378;vdb=variation;vf=105535077;sample=HG00109#373531_tablePanel >

Here we read this CSV file
```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.)/nrow(mxl)
```

Now let's look at a different population. I picked the GBR. 

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
table(gbr$Genotype..forward.strand.)/nrow(gbr)
```

This variant that is associated with childhood asthma is more frequent in
the GBR population than the MKL population.


Let's see if G|G versus A|G really affect the expression level of the gene
One sample is obviously not enough to know what is happening in a population. You are interested in assessing genetic differences on a population scale. So, you processed about ~230 samples and did the normalization on a genome level. Now, you want to find whether there is any association of the 4 asthma-associated SNPs (rs8067378...) on ORMDL3 expression.

>Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

How many samples do we have?
```{r}
expr <- read.table("expression_level.txt")
head(expr)
```
```{r}
table(expr$geno)
```

```{r}
nrow(expr)
```
```{r}
library(tidyverse)
expr %>%
  group_by(geno) %>%
  summarize(medianExp = median(exp))
```
 
 We have 462 samples including 108 for A/A, 233 for A/G and 121 for G/G. The median expression level for A/A is 31.24847, A/G is 25.06486, and G/G for 20.07363.  
 
>Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

Let's make a boxplot
```{r}
library(ggplot2)
ggplot(expr, aes(geno, exp, fill=geno))+
  geom_boxplot(notch=TRUE)+
  labs(x="Genotype", y="Normalized expression level", fill="Genotype")+
  theme_bw()
```


According to the boxplot, G/G is associated with lower gene expression levels compared to A/A. Therefore, we could conclude that this G/G SNP reduces the expression level of the gene *ORMDL3*.