---
title: "Halloween_Candy"
author: "Qianqian Tao"
format: html
---

```{r}
candy <- read.csv("candy-data.csv", row.names=1)

```

>Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

>Q2. How many fruity candy types are in the dataset?

```{r}
#using sum
sum(candy$fruity)
#using table
table(candy$fruity)

```
>Q2.5 What are these fruity candy?

```{r}
indices <- which(candy$fruity==1)

print_indices <- function(indices, dataset){
  for(i in indices){
    return(dataset[indices,])
  }
}
#print_indices(indices, candy)

#Another easier way to do this 
rownames(candy[candy$fruity==1,])

```

## How often does my favorite candy win 
>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Hershey's Krackel", ]$winpercent

candy["M&M", ]$winpercent

candy["Twix", ]$winpercent

```

>Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```

>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

```{r}
library("skimr")
#skimr::skim(candy)--> using only one function from the package without loading the whole package 
skim(candy)
```
>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Yeap, the `winpercent` column is on a 0:100 scale and all other appear to be 0: 1 scale. 

>Q7. What do you think a zero and one represent for the candy$chocolate column?

A 0 means this candy is not classified as containing chocolate and 1 means this candy is classified as containing chocolate.  

>Q8. Plot a histogram of winpercent values


```{r}
#one way to make a histogram in base R graphics: 
hist(candy$winpercent)

#use ggplot
library(ggplot2)

ggplot(candy)+
  geom_histogram(aes(winpercent), bins=10)
```

>Q9. Is the distribution of winpercent values symmetrical?

No. It is left skewed

>Q10. Is the center of the distribution above or below 50%?

It is below 50% with a mean:

```{r}
mean(candy$winpercent)

```

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

To answer this question I will need to :

-"subset" (a.k.a "select", "filter") the candy dataset to just chocolate candy,
- get their winpercent values,
-calculate the mean of these
Then do the same for fruity candy and compate.

```{r}
col_function <- function(dataset, col1, variable){
  mean(dataset[,col1][dataset[,variable]==1])
}
col_function(candy, "winpercent", "chocolate")

```



```{r}

chocolate_candy <- candy[candy$chocolate==1,]
choco_mean_win <- mean(chocolate_candy$winpercent)


#Professor's method
#Filter/select/subset to just chocolate rows
chocolate.candy <- candy[as.logical(candy$chocolate),]
#Get their winpercent values
chocolate.winpercent <- chocolate.candy$winpercent
#Calculate the mean value
mean(chocolate.winpercent)
```
```{r}

#Then do the same thing for fruit
#Filter/select/subset to just chocolate rows
fruity.candy <- candy[as.logical(candy$fruity),]
#Get their winpercent values
fruity.winpercent <- fruity.candy$winpercent
#Calculate the mean value
mean(fruity.winpercent)
```

Chocolate wins!

>Q12. Is this difference statistically significant?

```{r}
t.test(chocolate.winpercent, fruity.winpercent)
t.test(chocolate.candy, fruity.candy)
```
super low p-value--> there is significant difference between chocolate and fruit. 
>Q13. What are the five least liked candy types in this set?

```{r}
x <- c(5,2,10)
#use sort
sort(x)
#use order
order(x)
x[order(x)]
#I can order by winpercent
ord <- order(candy$winpercent)
head(candy[ord,],5)
```


```{r}

library(dplyr)
candy %>% arrange(winpercent) %>% head(5)
```

>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
#usual way
ord_2 <- order(candy$winpercent, decreasing=TRUE)
head(candy[ord_2,],5)
#one easy one
tail(candy[ord,],5)
```


```{r}

col_function <- function(dataset, col1, variable){
  mean(dataset[,col1][dataset[,variable]==1])
}
col_function(candy, "winpercent", "chocolate")

candy[,"winpercent"][candy[,"chocolate"]==1]
```
>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()+
  theme(axis.text.y = element_text(
          size = 3))
```
>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent))+
  geom_col()+
  theme(axis.text.y = element_text(
          size = 3))
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
# set color to my_cols only make the frame to be that color; set fill to my_cols make the filling color to be the corresponding my_cols
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) +
  theme(axis.text.y = element_text(
          size = 3))
```


> Q17. What is the worst ranked chocolate candy?

Sixlets

> Q18. What is the best ranked fruity candy?

Starburst

> Q. What is the best candy 

```{r}
library(ggrepel)
my_cols[as.logical(candy$fruity)] = "blue"
ggplot(candy, aes(winpercent, pricepercent))+
  geom_point(col=my_cols)
```

Add some labels 
```{r}

ggplot(candy, aes(winpercent, pricepercent, label=rownames(candy)))+
  geom_point(col=my_cols)+
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 10)

```
>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Reese's Miniature or Reese's Peanut Butter cup (chocolate)

```{r}
order_win <- order(candy$winpercent)
tail(candy[order_win,])
```


>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?


```{r}
order_price <- order(candy$pricepercent)
tail(candy[order_price,])
```

# 5 Exploring th ecorrelations structure

Pearson correlation goes between -1 and +1 with zero indicating no correlation and values close to 1 being highly correlated.
```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```

>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

fruit and chocolate are anti-correlated
>Q23. Similarly, what two variables are most positively correlated?

chocolate and winpercent or chocolate and bar are most positively correlated

#Principal Component Analysis

The base R function for PCA is called `pcromp()` and we can set "scale=TRUE/FALSE"
```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```
The main result of PCA - i.e. the new PC plot (projection of candy on our new PC axis) is contained in `pca$x`
```{r}
pc <- as.data.frame(pca$x)
ggplot(pc)+
  aes(PC1, PC2, label=rownames(pc))+
  geom_point(col=my_cols)
 # geom_text_repel(max.overlaps = 10)

```
```{r}
ggplot(pc)+
  aes(PC1, PC2, label=rownames(pc))+
  geom_point(col=my_cols)+
  geom_text_repel(max.overlaps = 10)
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, hard and pluribus 

PC 1 captures correlation structure. If a candy is fruity, hard and pluribus, it will be on the positive side of the axis. 