---
title: "Lab8"
author: "Qianqian Tao"
format: pdf
---

## Quarto
Today we are going to explore some data from the University of Wisconsin Cancer Center on Breast biopsy data. 
```{r}
wisc.df <- read.csv("WisconsinCancer.csv", row.names=1)
head(wisc.df)

```
>Q1. How many observations are in this dataset?

```{r}
nrow(wisc.df)
```

There are `r nrow(wisc.df)` patients in this dataset.

>Q2. How many of the observations have a malignant diagnosis?

```{r}
diagnosis <- wisc.df$diagnosis
diagnosis_factor <- as.factor(diagnosis)
table(diagnosis_factor)
```
Now exclude the diagnosis column from the data
```{r}
wisc.data <- wisc.df[,-1]
```
>Q. How many "dimensions", "variable", "columns" are there in this dataset?

```{r}
ncol(wisc.data)
```

>Q3. How many variables/features in the data are suffixed with _mean?

```{r}
sum(grepl("_mean",colnames(wisc.data)))

```

# Principal Component Analysis (PCA)

To perform PCA in R we can use the `prcomp()` function. It takes an input as a numeric dataset and optional `scale=FALSE/TRUE` argument.

We generally always want to set `scale=TRUE` but let's make sure by checking if the mean and standard deviation values are different across these 30 columns.
```{r}
round(colMeans(wisc.data))

apply(wisc.data,2,sd)
```
>Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

PC1 captures 44.27% of variance
> Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

We need PC1, PC2 and PC3 to describe at least 70% of the variance.

> Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

We need PC1 to PC7 to describe 90%. 

```{r}

wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```
```{r}
attributes(wisc.pr)
plot(wisc.pr$x[,1],wisc.pr$x[,2],col=diagnosis_factor,xlab = "PC1", ylab = "PC2" )

```
```{r}
plot(wisc.pr$x[, 1],wisc.pr$x[, 3],col = diagnosis_factor, 
     xlab = "PC1", ylab = "PC3")
```
> Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?

It shows which variables are driving the differences, but it is hard to understand because the names are numbers. 

```{r}
biplot(wisc.pr)
```


```{r}
library(ggplot2)
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis_factor

ggplot(df, aes(x=PC1, y=PC2, col=diagnosis))+
  geom_point()
```
>Q. How much variance is captured in the top 3 PCs?

They capture 72.636% of the total variance. 

> Q9. For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean?

```{r}
wisc.pr$rotation["concave.points_mean",1]
```

```{r}
attributes(wisc.pr)
```
# Combine PCA results with clustering.

We can use our new PCR variables (i.e. the scores along the PCs contained in `pca$x` ) as input for other methods such as clustering.

```{r}
#Hclust needs a distance matrix as input
d <- dist(wisc.pr$x[, 1:3])

hc <- hclust(d, method = "ward.D2")
plot(hc)
```
To get our cluster membership vector we can use the `cutree()` function and specify a height (`h=`) or number of groups (`k`).

```{r}
grps <- cutree(hc, h=80)
table(grps)
```

I want to find out how many diagnosis "M" and "B" are in each grp?

```{r}
table(diagnosis,grps)
```
We can also plot our result using our clustering vector `grps`
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps)
```

```{r}
ggplot(df, aes(x=PC1, y=PC2, col=grps))+
  geom_point()
```
>Q15. What is the specificity and sensitivity of our current results?

```{r}
#Sensitivity
sensitivity <- 179/(179+33)

#Specificity
specificity <- 333/(333+24)

```
The sensitivity is `r sensitivity` and specificity is `r specificity`.

#Prediction

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc

```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```

> Q16. Which of these new patients should we prioritize for follow up based on your results?



