---
title: "Lab6_supplement"
author: "Qianqian Tao"
format: pdf
---

```{r}
# (A. Can you improve this analysis code?
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
df_unmodified <- df
df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a))
df$b <- (df$b - min(df$a)) / (max(df$b) - min(df$b))
df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c))
df$d <- (df$d - min(df$d)) / (max(df$a) - min(df$d)) 

```

```{r}
# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug

s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

>Q1.

read.pdb() reads a Protein Data Bank (PDB) coordinate file and returns a list of class "pdb" with 8 components.

>Q2.

trim.pdb() trims a pdb object to a subset of atoms.

>Q3. 

Deleting "sse=s1.chainA" would turn off the marginal black and grey rectangles. "sse" represents secondary structure object as returned from read.pdb.
```{r}
plotb3(s1.b, typ="l", ylab="Bfactor")
```
> Q4.

A plot that combines the above plots together but shows them in different colors. 

> Q5. 

Protein 1 and protein 3 are more similar to each other.

```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

>Q6.

```{r}
#This is a function that takes a pdb file name as input and generate a plot as the output.
analysis_code <- function(pdb){
  #Input a pdb file name with quotation marks, e.g. pdb <- "4AKE"
  protein <- read.pdb(pdb)
  protein.chainA <- trim.pdb(protein, chain="A", elety="CA")
  protein.b <- protein.chainA$atom$b
  #Generate a plot as the output
  plotb3(protein.b, sse=protein.chainA, typ="l", ylab="Bfactor")
}

analysis_code("4AKE")
analysis_code("1AKE")
analysis_code("1E4Y")


```

