---
title: "Lab6"
author: "Qianqian Tao"
format: pdf
---

Q1. Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an NA value) this can be used as a score to be potentially dropped. Your final function should be adquately explained with code comments and be able to work on an example class gradebook such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]

All functions in R have at least 3 things:

-**Name**, in our case"grade"
-Input **argument**, student1 etc.
-**Body**, this is our working snippet above

```{r}

student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)

x <- student3
x[is.na(x)] <- 0
mean(x)

grade <- function(vector){
  #Map/Replace NA values to zero
  vector[is.na(vector)] <- 0
  #min_index <- which.min(vector)
  #Exclude the lowest score
  #vector <- vector[-which.min(vector)]
  #Exclude the lowest score and calculate the mean 
  mean(vector[-which.min(vector)])
}

grade(student1)
grade(student2)
grade(student3)
x
min(x)
which.min(x)
```
Q2. Using your grade() function and the supplied gradebook, Who is the top scoring student overall in the gradebook? [3pts]
```{r}
#read.csv("https://tinyurl.com/gradeinput")
class_data <- read.csv("https://tinyurl.com/gradeinput", row.names=1)
class_average <- apply(class_data,1,grade)
which.max(apply(class_data,1,grade))
class_average[which.max(class_average)]

```

Q3. From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall? [2pts]
```{r}
new_mean <- function(vector){
  vector[is.na(vector)] <- 0
  mean(vector)
}
hw_average <- apply(class_data, 2, new_mean)
hw_average[which.min(apply(class_data, 2, new_mean))]
which.min(apply(class_data, 2, new_mean))


```

Q4. Optional Extension: From your analysis of the gradebook, which homework was most predictive of overall score (i.e. highest correlation with average grade score)? [1pt]
```{r}
cor(class_data$hw1,class_average)
cor(class_data$hw2,class_average)
cor(class_data$hw3,class_average)
cor(class_data$hw4,class_average)
```

If I try on hw2, I get NA as there are missing homeworks. I will mask all NA values to Zero
```{r}
mask <- class_data
mask[is.na(mask)] <- 0
mask
cor(mask$hw1,class_average)
cor(mask$hw2,class_average)
cor_homework <- apply(mask, 2, cor, y=class_average)
cor_homework[which.max(apply(mask, 2, cor, y=class_average))]


```

